---
title: "Project Draft"
author: "Ria Pinjani"
date: "10/19/2019"
output: html_document
---

Setup

```{r ,include=FALSE}
setwd('/Users/riapinjani/Documents/Fall\'19/Statistical\ Consulting')
```


```{r}

library('readxl')
require('plyr')
require('sas7bdat')
require('tidyr')
require('sf')
require('mapview')
require('lubridate')
require('ggplot2')
require('tidyr')
require('ggmap')
require('tigris')
```

Read in tobacco retailers data set for 2007 - 2015

```{r}

foil_2007_2015  <- read_excel('retailers data/FOIL Case No. 83-2017_List of Cigarette Retail Dealers from 2007 to 2015.xlsx')
```

Create variable year from CREATE_DATE

```{r}
foil_2007_2015 <-  foil_2007_2015  %>% mutate (Year = year(foil_2007_2015$CREATE_DATE)) 
```


Subset to years 2010 - 2015

```{r}
foil_2010_2015 <- subset(foil_2007_2015, Year %in% c(2010,2011,2012,2013,2014,2015))
```

Number of retailers for each year from 2010 - 2015

```{r}
foil_2010_2015_n <- foil_2010_2015 %>% count('Year')
foil_2010_2015_n
```





```{r}
ggplot(foil_2010_2015_n %>% count('Year'), aes ( x = Year, freq)) + geom_line() + theme_bw() + ylab('Count') + ggtitle('Retailers by year')
```

```{r}
ggsave('Retailers by year.png')
```

Data Cleaning - Remove values where BORO = outside NYC or NULL and values where STATE = NULL

```{r}
table(foil_2010_2015$BORO)
table(foil_2010_2015$STATE)
```



```{r}
foil_2010_2015$BORO <- toupper(foil_2010_2015$BORO)
foil_2010_2015 <- subset(foil_2010_2015 , BORO %in% c('BRONX', 'BROOKLYN', 'MANHATTAN', 'QUEENS', 'STATEN ISLAND'))
```


```{r}
foil_2010_2015 <- subset(foil_2010_2015, STATE == 'NY')
```

Make characters in the column CITY to be consistent

```{r}
foil_2010_2015$CITY[which(foil_2010_2015$CITY == "BAYSIDE HILLS")] = "BAYSIDE"
foil_2010_2015$CITY[which(foil_2010_2015$CITY == "CAMBRIA HTS")] = "CAMBRIA HEIGHTS"
```

Create column address

```{r}
foil_2010_2015 <- unite(foil_2010_2015,"address", c(BUILDING, STREET, CITY, STATE, ZIP), sep = " ")
```

Remove column STREET_2 as it does not hold any information 

```{r}
foil_2010_2015$STREET_2 <- NULL
```

Create column Year2 from EXPIRATION_DATE

```{r, warning=FALSE}
foil_2010_2015 <-  foil_2010_2015  %>% mutate (Year2 = year(foil_2010_2015$EXPIRATION_DATE)) 
```

Data set for the year 2011

```{r}
foil_2011 <- subset(foil_2010_2015, Year == 2011)
```

Google API for geocoding addresses


```{r, warning=FALSE}
register_google(key = 'AIzaSyA8y8FgxCH4WamYldyyBZyIg1TFUTLGUo0')
```

Geocoding for the year 2011

Manhattan

```{r}
manhattan2011 <- subset(foil_2011, BORO == 'MANHATTAN')
```

```{r, include=FALSE}
geocode_manhattan2011 <- geocode(manhattan2011$address)
```


Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_manhattan2011$census_code <- apply(geocode_manhattan2011, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```

```{r}
geocode_manhattan2011$census_tract <- substr(geocode_manhattan2011$census_code, 1, 11)
```



```{r}
manhattan2011 <- as.data.frame(cbind(manhattan2011, geocode_manhattan2011))
```

Brooklyn


```{r}
brooklyn2011 <- subset(foil_2011, BORO == 'BROOKLYN')
```


```{r, include=FALSE}
geocode_brooklyn2011 <- geocode(brooklyn2011$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame


```{r}
geocode_brooklyn2011$census_code <- apply(geocode_brooklyn2011, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```

```{r}
geocode_brooklyn2011$census_tract <- substr(geocode_brooklyn2011$census_code, 1, 11)
```


```{r}
brooklyn2011 <- as.data.frame(cbind(brooklyn2011, geocode_brooklyn2011))
```

Bronx

```{r}
bronx2011 <- subset(foil_2011, BORO == 'BRONX')
```


```{r, include=FALSE}
gecode_bronx2011 <- geocode(bronx2011$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame


```{r}

gecode_bronx2011$census_code <- apply(gecode_bronx2011, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```


```{r}
gecode_bronx2011$census_tract <- substr(gecode_bronx2011$census_code, 1, 11)
```


```{r}
bronx2011 <- as.data.frame(cbind(bronx2011, gecode_bronx2011))
```

Queens

```{r}
queens2011 <- subset(foil_2011, BORO == 'QUEENS')
```


```{r, include=FALSE}
geocode_queens2011 <- geocode(queens2011$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_queens2011$census_code <- apply(geocode_queens2011, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```



```{r}
geocode_queens2011$census_tract <- substr(geocode_queens2011$census_code, 1, 11)
```


```{r}
queens2011 <- as.data.frame(cbind(queens2011, geocode_queens2011))
```


Staten Island

```{r}
statenisland2011 <- subset(foil_2011, BORO == 'STATEN ISLAND')
```


```{r, include=FALSE}
geocode_statenisland2011 <- geocode(statenisland2011$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_statenisland2011$census_code <- apply(geocode_statenisland2011, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```


```{r}
geocode_statenisland2011$census_tract <- substr(geocode_statenisland2011$census_code, 1, 11)
```


```{r}
statenisland2011 <- as.data.frame(cbind(statenisland2011, geocode_statenisland2011))
```

Create data frame for ny 2011 combine all 5 boroughs

```{r}
ny2011 <- as.data.frame(rbind(manhattan2011, brooklyn2011, bronx2011, queens2011, statenisland2011))
```





Data set for the year 2015

```{r}
foil_2015 <- subset(foil_2010_2015, Year == 2015)
```

Geocoding for the year 2015

Manhattan


```{r}
manhattan2015 <- subset(foil_2015, BORO == 'MANHATTAN')
```

```{r, include=FALSE}
geocode_manhattan2015 <- geocode(manhattan2015$address)
```


Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_manhattan2015$census_code <- apply(geocode_manhattan2015, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```


```{r}
geocode_manhattan2015$census_tract <- substr(geocode_manhattan2015$census_code, 1, 11)
```


```{r}
manhattan2015 <- as.data.frame(cbind(manhattan2015, geocode_manhattan2015))
```

Brooklyn


```{r}
brooklyn2015 <- subset(foil_2015, BORO == 'BROOKLYN')
```


```{r, include=FALSE}
geocode_brooklyn2015 <- geocode(brooklyn2015$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_brooklyn2015$census_code <- apply(geocode_brooklyn2015, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```

```{r}
geocode_brooklyn2015$census_tract <- substr(geocode_brooklyn2015$census_code, 1, 11)
```

```{r}
brooklyn2015 <- as.data.frame(cbind(brooklyn2015, geocode_brooklyn2015))
```

Bronx

```{r}
bronx2015 <- subset(foil_2015, BORO == 'BRONX')
```


```{r, include=FALSE}
gecode_bronx2015 <- geocode(bronx2015$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

gecode_bronx2015$census_code <- apply(gecode_bronx2015, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```


```{r}
gecode_bronx2015$census_tract <- substr(gecode_bronx2015$census_code, 1, 11)
```


```{r}
bronx2015 <- as.data.frame(cbind(bronx2015, gecode_bronx2015))
```


Queens

```{r}
queens2015 <- subset(foil_2015, BORO == 'QUEENS')
```


```{r, include=FALSE}
geocode_queens2015 <- geocode(queens2015$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_queens2015$census_code <- apply(geocode_queens2015, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```

```{r}
geocode_queens2015$census_tract <- substr(geocode_queens2015$census_code, 1, 11)
```

```{r}
queens2015 <- as.data.frame(cbind(queens2015, geocode_queens2015))
```

Staten Island

```{r}
statenisland2015 <- subset(foil_2015, BORO == 'STATEN ISLAND')
```


```{r, include=FALSE}
geocode_statenisland2015 <- geocode(statenisland2015$address)
```

Using the tigris package to identify cencus tract for each tobacco retailers from lat and lon variables & Combine into one data frame

```{r}

geocode_statenisland2015$census_code <- apply(geocode_statenisland2015, 1, function(row) call_geolocator_latlon(row['lat'], row['lon']))

```

```{r}
geocode_statenisland2015$census_tract <- substr(geocode_statenisland2015$census_code, 1, 11)
```

```{r}
statenisland2015 <- as.data.frame(cbind(statenisland2015, geocode_statenisland2015))
```

Create data frame for ny 2015 combine all 5 boroughs


```{r}
ny2015 <- as.data.frame(rbind(manhattan2015, brooklyn2015, bronx2015, queens2015, statenisland2015))
```


Write csv files for 2011 and 2015 data sets

```{r}


write_csv(ny2011, "ny2011.csv")
write_csv(ny2015, "ny2015.csv")

```























































